# Exploratory Data Analysis - Raw Data

The scripts below are exploratory in nature and provide (insight in):

 - missingness and patterns of missingness in the data
 - distrutions of values for all variables in the data
 - some quantitative measures of variables and descriptive statistics
 - some exploratory plots

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  error = FALSE,
  message = FALSE)
```

## Packages
The packages needed for this analysis.
```{r}
library(tidymodels)
library(tidyverse)
library(psfmi)
library(miceadds)
library(micemd)
library(mitools)
library(mice)
library(naniar)
library(skimr)
#library(norm)
## warning
## This package has some major limitations
#(for example, it does not work reliably when
#  the number of variables exceeds 30),
#and has been superseded by the norm2 package.
#detach("package:VIM", unload = TRUE)
# I replaced the norm for norm2 package because of the above warning.
#library(norm2)
library(VIM)
library(finalfit)
library(haven)
```

## Load data
The data is loaded from the original data file, which is an SPSS `.sav` file. The package `{haven}` is used to parse the file in R.
```{r}
df_raw <-
  read_sav(
    here::here(
      "data-raw",
      "D010",
      "SPSS basis File results.sav"
      )
    ) |> 
  janitor::clean_names()

```

## First glimpse at missingness
Let's explore the missing values first with a visual.
```{r}
naniar::vis_miss(df_raw)

df_raw <- df_raw |>
  mutate(index = 1:nrow(df_raw))

```


## Select relevant variables
To select only the variables we need, we make a dataframe containing only those.
```{r}
df_select <- df_raw |>
   dplyr::select(-c(
    # patient_code,
     attitude,
     pt_code,
     coping_active,
     coping_passive,
     percentage_active_coping,
     percentage_passive_coping,
     pain_6weeks,
     pain_categorized_6weeks,
     pain_3months,
     pain_catagorized_3months,
     pain_6months,
     pain_categorized_6months,
     index)
     ) |>
  rename(sex = seks)

df_select
names(df_select)
#DT::datatable(df_select)
```

## Exploratory Data Analysis
In the section below, the data is explored by diving more into detail on the missingness, the data distributions and some descriptive statistics.

### Exploratory plots
First, we take a look at the data using a number of different graphical representations.
```{r}
## males / females
df_select |>
  group_by(sex) |>
  tally() |>
  ggplot(aes(x = as_factor(sex), y = n)) +
  geom_col(show.legend = FALSE)

## age
df_select |>
  group_by(sex) |>
  summarise(mean_age = mean(age, na.rm = TRUE)) |>
  ggplot(aes(x = as_factor(sex), y = mean_age)) +
  geom_col(show.legend = FALSE)

## boxplot  
df_select |>
  group_by(sex) |>
  ggplot(aes(x = as_factor(sex), y = age)) +
  geom_boxplot(show.legend = FALSE)

## age distribution
df_select |>
  group_by(sex) |>
  ggplot(aes(x = age)) +
  geom_density(aes(colour = as_factor(sex))) +
  theme_bw()

## pain intensity
df_select |>
  group_by(sex) |>
  ggplot(aes(x = pain_intensity)) +
  geom_density(aes(colour = as_factor(sex)), size = 1) +
  theme_bw()

## dotplots
df_select |>
  group_by(sex, smoking) |>
  ggplot(aes(x = as_factor(sex), y = disability)) +
  geom_point(
    aes(colour = as_factor(smoking)), 
    position = "jitter", 
    shape = 1, 
    alpha = 0.8) +
  theme_bw() +
  xlab(NULL)

## work
df_select |>
  group_by(work) |>
  tally()

```

### Missing values
Some overall missingness inspections. Here we explore the numerical characteristics of the missingness of the data.
```{r}
# Are there missing values in the dataset?
any_na(df_select)
# How many?
n_miss(df_select)
prop_miss(df_select)
# Which variables are affected?
df_select %>% 
  is.na() %>% 
  colSums() |> 
  enframe() |> 
  ggplot(aes(x = reorder(as_factor(name), value), y = value)) +
    geom_point() +
    coord_flip() +
  xlab("variable in data") +
  ylab("Total number of NA")

```

### Get number of missings per variable (n and %)
```{r}
miss_var_summary(df_select)
miss_var_table(df_select)
# Get number of missings per observation (n and %)
miss_case_summary(df_select)
miss_case_table(df_select)
```

### Some more plots on missingness
This plot shows the number of missing values per individual variable.
```{r}
gg_miss_var(df_select)
```

With this plot we can judge whether missingness is at random (or not). We see no obvious pattern in the missinness.
```{r}
naniar::vis_miss(df_select) + 
  theme(axis.text.x = element_text(angle=80)) +
  coord_flip()
```

Is the presence of missing values related with missings in other variables?
```{r}
gg_miss_upset(df_select)
```

From this plot we see that much of the missingness is not accross many variables. There are however quite a number of cases (88) where missing cases are accross the vars `painint_3months_NA`, `painint_6months_NA` and `painint_6 weeks_NA`. This suggests some missingness not at randomm accross these variables and warrants a closer inspection. 

From the [blog](https://towardsdatascience.com/smart-handling-of-missing-data-in-r-6425f8a559f2)

If you had concrete hypothesis about the impact of the presence of missing values in a certain variable on a target variable, you could test it like this:

```{r}
# add a variable to the dataset that indicates the missingness of paininit_6weeks per observation
df_select_test <- df_select %>%
  mutate(missing_6weekspainit = is.na(painint_6weeks))
  
# get missing paininit_6weeks info for participants that have a therapist with "Biopsychosocial (value 2)" attitude
missing_paininit_6weeks_male <- df_select_test %>%
  filter(sex == 1) %>%
  pull(missing_6weekspainit)
  
# get missing paininit_6weeks info for participants that have a therapist with "Biomedical (value 1)" attitude
missing_paininit_6weeks_female <- df_select_test %>%
  filter(sex == 2) %>%
  pull(missing_6weekspainit)
  
#check whether the percentage of missings in missing_paininit_6weeks differ per level of sex.
t.test(
  missing_paininit_6weeks_male, 
  missing_paininit_6weeks_female
  )
```
There is no reason to assume (from this analysis) that the sex of a patient is correlated to having missing values for paininit_6weeks. Note that this is a hypothetical case, that might not make sense in this particular case, when you bring context into play.

## Write table with all labels
The `.sav` contains so-called SPSS labels that hold information about how values for dichotmous and discrete variables are encoded. WE can use this information by creating a table and use this table to recode some of the variables. 
```{r}
data_label <- t(as.data.frame(labelled::var_label(df_select))) 
data_label

df = df_select
ind = 2
## get labels and coding
get_spss_coding <- function(ind, df){
  
  at <- attributes(df[[ind]])$labels |> enframe()
  at <- map_df(
    .x = at,
    .f = as.character
  )
  ## get column name
  col <- names(df[ind])
  at$var_name <- col |> as.character()
  
  ## return df
  return(at)
  
  }

map_df(
  .x = 1:ncol(df_select),
  .f = get_spss_coding,
  df = df_select
) -> labelling
labelling

DT::datatable(labelling)

## remove sps-labels
df_select <- df_select |>
  sjlabelled::remove_all_labels()

```


## Deal with 'work' variables
The variable 'work' need special attention. It's value depends on some other work-related variables.

We recode the 'work' variable as follows:
Work 1 = yes / 2 = no
When variable 'work' = 2, recode 'work_happiness', 'work_satisfaction' en 'posture_work' to value: 3


```{r}
df_select |>
  dplyr::select(
    work,
    work_happiness,
    work_satisfaction,
    posture_work
  ) -> df_work

map_df(
  .x = df_work,
  unique
)

df_work |>
  dplyr::filter(work == 2) |> naniar::vis_miss()


df_select$work |> unique()
df_select$work_happiness |> unique()
df_select$work_satisfaction |> unique()
df_select$posture_work |> unique()


df_select <- df_select |>
  mutate(
    work_happiness = ifelse(
      work == 2, 3, work_happiness 
    )
  )

df_select <- df_select |>
  mutate(
    work_satisfaction = ifelse(
      work == 2, 3, work_satisfaction
    )
  )

df_select <- df_select |>
  mutate(
    posture_work = ifelse(
      work == 2, 3, posture_work 
    )
  )

## Check if NA's are replaced for "NOT-APPLICABLE
df_work |>
  dplyr::filter(work == 2) |> naniar::vis_miss()

df_select |>
  ggplot(
    aes(
      x = as_factor(work),
      y = as_factor(work_happiness)
    )
  ) +
  geom_point(position = "jitter", colour = "darkblue") +
  toolboxr::rotate_axis_labels("x", angle = 90) -> work_plot_happy

work_plot_happy

df_select |>
  ggplot(
    aes(
      x = as_factor(work),
      y = as_factor(work_satisfaction)
    )
  ) +
  geom_point(position = "jitter", colour = "darkblue") +
  toolboxr::rotate_axis_labels("x", angle = 90) -> work_plot_satisfaction
work_plot_satisfaction

df_select |>
  ggplot(
    aes(
      x = as_factor(work),
      y = as_factor(posture_work)
    )
  ) +
  geom_point(position = "jitter", colour = "darkblue") +
  toolboxr::rotate_axis_labels("x", angle = 90) -> work_plot_posture
work_plot_posture

cowplot::plot_grid(
  work_plot_happy,
  work_plot_posture,
  work_plot_satisfaction,
  ncol = 1
)


```

## Recode `physical_activity`
The variable 'physical_activity' has 3 levels. In order to transform this into a dichotmous variable (with only two outcomes), we need some transformations.

1 = I do not move on one single day of the week
2 = I am active on 5 or more days a week
3 = I am in between score 1 and 2
Score 1 and 3 need to be recoded to value 1  (Not achieving the Dutch Healthy Exercise Norm)
And score 2 needs to be recoded to 0  (Achieving the Dutch Healthy Exercise Norm )

```{r}
df_select$physical_activity |> unique()

df_select <- df_select |>
  mutate(
    physical_activity_recoded = 
      ifelse(
        (physical_activity == 1 | physical_activity == 3),
        1, 
        physical_activity),
    physical_activity_recoded = 
      ifelse(
        physical_activity == 2, 
        0, 
        physical_activity_recoded
      
    ) 
  )

df_select |>
  ggplot(
    aes(
      x = as_factor(physical_activity),
      y = as_factor(physical_activity_recoded)
    )
  ) +
  geom_point(position = "jitter", colour = "darkblue") +
  toolboxr::rotate_axis_labels("x", angle = 20)


df_select$physical_activity <- df_select$physical_activity_recoded
df_select <- df_select |>
  dplyr::select(-physical_activity_recoded)

```

## Write subsetted data to disk
As a final step we write the cleaned up subsetted file to disk for down the line use.
```{r}

write_rds(
  df_select,
  here::here(
    "data",
    "df_non_imp_select.rds"
  )
)

## and the labelling
write_csv(
  labelling,
  here::here(
    "data",
    "labelling_and_coding_vars.csv"
  )
)
```
