# Imputation of missing values

This section describes in detail how we dealt with imputation of missing values in the data. The overal method of imputation that was chosen is the multiple imputation methods, available in the `{mice}` R package.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r, packages}
library(tidyverse)
library(mice)
```

## Data
Because we need to work with multiple versions of the dataset, I decided to change the data structure into a list. This makes it more traceable and more easy to inspect different datasets and keep track of changes. At the end of this analysis a single list object will have all the different versions of the datasets collected together and written to disk.

```{r, load_data}
## raw data
df_raw <- haven::read_sav(
  here::here(
    "data-raw",
    "D010",
    "SPSS basis File results.sav"
  )
)

## non-imputed - subset
df_non_imp_select <- readr::read_rds(
  here::here(
    "data",
    "df_non_imp_select.rds"
  )) |> as_tibble()

names(df_non_imp_select)

data <- list(
  data_raw = df_raw,
  data_subset = df_non_imp_select
)

rm(df_raw, df_non_imp_select)

## add metadata
attr(data$data_subset, "meta") <- "Data variable subset from raw data [file: './data-raw/D010/SPSS basis File results.sav']. Dataframe stored on disk as [file: './data/df_non_imp_select.rds']"  

attr(data$data_raw, "meta") <- "Raw unprocessed data as recieved from supplier [file: './data-raw/D010/SPSS basis File results.sav']" 

## get attributes
attributes(data$data_raw)
```

## Prepare datset for imputing
To prepare the data for imputation we need to carefully consider which variables to keep in the data and which variables we will use to get imputed values for missing values. In order to have a dataset that we can staert preparing, we copy the dataset so that we can keep the original data for comparison.

```{r}
## copy datasets
data$data_to_impute <- data$data_subset
names(data)

data_to_impute <- data$data_to_impute
names(data_to_impute)



```

The id column 'patient_code' is a zero variance variable and does not serve any pupose in the imputation process, so here we delete it.
```{r}
## remove ID var 'patient_code'
data$data_to_impute <- data$data_to_impute |>
  select(-patient_code)

```

## Convert all categorical vars to factors
For the imputation process we need the discrete variables to be factors.
```{r, data_distributions}
## which are factors and which are continuous vars?
should_be_factors <- c(
  "coping",
  "sleep_quality",
  "alcohol",
  "smoking",
  "physical_activity",
  "posture_work",
  "work_satisfaction",
  "work_happiness",
  "work",
  "headache",
  "wide_spread_pain",
  "recurrence",
  "education_level",
  "education",
  "sex"
)

## copy dataset
data$data_preprocessed <- data$data_to_impute

## convert the categoricals to factors
data$data_preprocessed <- data$data_preprocessed %>%
  mutate(across(should_be_factors, as_factor))

## check datatypes
map_df(
  .x = data$data_preprocessed,
  .f = class
  ) |> 
  t() |>
  enframe(name = "var_name", value = "var_type") -> var_types
```

## Panel with all distributions
In a panel plot we can show all the distibutions which provide an idea on which variables are discrete and which are continuous. The panels can also be viewed as images in the `./img` folder.
```{r, distributions_panel, fig.height=30, fig.width=30}
##  function to plot a vectors histogram
source(
  here::here(
    "R",
    "plot_distro.R"
  )
)

ind_num <- map_lgl(
  .x = data$data_preprocessed,
  .f = is.numeric
)

num_vars <- names(data$data_preprocessed)[ind_num]

## create histogram to establish var type / data distribution
list_histograms <- map2(
  .x = data$data_preprocessed[ ,ind_num],
  .y = num_vars,
  .f = plot_distro
)

list_histograms

cowplot::plot_grid(plotlist = list_histograms)
ggsave(here::here("img", "distros.png"), height = 40, width = 40, dpi = 300)
ggsave(here::here("img", "distros.svg"), height = 40, width = 40, dpi = 300)
```

# Imputation of missing values

## Checking Missing Completely at Random (MCAR)
https://bookdown.org/mwheymans/bookmi/missing-data-evaluation.html#missing-data-evaluation-in-r

From the analysis above we can conclude that some of the variables:
are on a continuous or discrete scale. We need to select those to assess MCAR.

### Missingness pattern
```{r, eval=FALSE}
md.pattern(data$data_preprocessed) ## figure does not scale properly, TODO: needs a fix
```

## Define predictors to include in the imputations
```{r, selecting_numeric_vars}

## select continuous and discrete
var_selection <- c(
  # "patient_code", 
  "sex",
  "age",
  "education",
  "education_level",
  "pain_intensity",
  "duration",
  "recurrence",
  "wide_spread_pain",
  "headache",
  "disability",
  "work",
# "work_happiness",
# "work_satisfaction",
# "posture_work",
  "physical_activity",
  "smoking",
  "alcohol",
  "bmi",
  "sleep_quality",
  "catastrophizing",
  "duration_beliefs",
  "concerns",
  "treatment_beliefs",
  "depression",
  "kinesiophobia",
  "distress",
  "coping",
  "identity_beliefs",
  "hypervigilance",
  "self_efficacy",
  "relation",
  "painint_6weeks",
  "painint_3months",
  "painint_6months"
  )

intersect(names(data$data_preprocessed), var_selection)

library(naniar)
mcar_test(data = data$data_preprocessed[, var_selection])
```

From `?mcar_test`:
Use Little's (1988) test statistic to assess if data is missing completely at random (MCAR). The null hypothesis in this test is that the data is MCAR, and the test statistic is a chi-squared value. 

Based on the fact the p-value here is not < 0.05, we can conclude that we cannot reject the NULL hypothesis, thus concluding that the missingness in the data is MCAR.

## Inspect the unique values of some of the variables
```{r, unique_values, eval=FALSE}
df_select_compare <- data$data_preprocessed |>
  dplyr::select(
    age,
    depression,
    duration,
    pain_intensity
  )

map(
  .x = df_select_compare, 
  .f = unique
  )


df_select_compare_fct <- df_select_compare |>
  mutate(
    age = as_factor(age),
    depression = as_factor(depression),
    duration = as_factor(duration)
  )
  
  
map(
  .x = df_select_compare_fct, 
  .f = levels
  )

map(
  .x = df_select_compare_fct, 
  .f = unique
  )

rm(df_select_compare, df_select_compare_fct)

```

## Using the MICE package for imputation of missing values
Before we proceed with MICE, we need to remove the SPSS labels from the dataframe (see: https://stackoverflow.com/questions/62395629/error-t-haven-labelled-not-supported-while-attempting-mice-package-in-r)

## Create predictorMatrix for MICE

When we want to create a predictionMatrix automatically with a threshold for low correlating variables

```{r, preditor_matrix_imp}
names(data$data_preprocessed)
pred <- quickpred(data$data_preprocessed, minpuc = 0.5, include = var_selection)

rownames(pred) <- colnames(pred)
pred <- pred |> as.matrix()
pheatmap::pheatmap(pred, cluster_rows = FALSE, cluster_cols = FALSE)

```

## Calculate percentage missing data and cases
When we want to have all variables in the dataset available for imputation

One of the hyperperameters for the 'mice' function is `m`, the number of imputations: this should be at least the percentage of missing cases in the data.

Let's calulate

```{r}
## manual calculation
n_na <- sum(is.na(data$data_preprocessed))
n_obs <- nrow(data$data_preprocessed) * ncol(data$data_preprocessed)

perc_mssing <- 100 * (n_na/n_obs)
perc_mssing |> round(1)

## using naniar package / visual confirmation is always good
vis_miss(data$data_preprocessed)

## missing cases
pct_missing_cases <- naniar::prop_miss_case(data$data_preprocessed) * 100
```

So `r perc_mssing`% of the total data is missing. But `r pct_missing_cases`% of the cases has at least one missing value. We set the `m` parameter to 75   

## Running the imputations
In order to allow for running imputations over multiple cores of the computer, we use the `{future}` package.

```{r, run_imputations, cache=TRUE}
df_inspect <- data$data_preprocessed

## get data types
map_df(
  df_inspect,
  class
) |> 
  t() |>
  as_tibble() -> datatypes


levels(df_inspect$physical_activity)
df_inspect$physical_activity |> unique()

levels(df_inspect$work)
df_inspect$work |> unique()

levels(df_inspect$work_happiness)
df_inspect$work_happiness |> unique()

levels(df_inspect$posture_work)
df_inspect$posture_work |> unique()


future::plan("multisession", workers = pmin(2L, future::availableCores()))
data$mult_imp <- futuremice(
  data$data_preprocessed, 
  m = 75,
  maxiter = 10,
  parallelseed = 123,
  predictorMatrix = pred
  )
```

## Inspect the imputations
We can look at the imputed dataframe as an item in the 'data' list object.
```{r}
data$mult_imp
```

Printing the `mult_imp` object to the console will give you the used methods for each variable:

**It is suggested to let the MICE function decide the method and not change it.**
see also: https://stats.stackexchange.com/questions/204313/how-to-choose-which-imputation-to-use-to-replace-missing-values

## Check convergence
```{r}
plot(data$mult_imp)
```

From these plots we see that there is no evident trend in the estimation of the means over iterations, for any of the imputed variables. Therefore we can conclude that convergence was achieved and that the max. iteration of 10 is a valid setting for this imputation run.

## Check for plausible values of imputation
```{r}
svg(filename = here::here("img", "stripplot.svg"))
stripplot(data$mult_imp)
dev.off()
```

## Checking the used predictor matrix
```{r}
future::plan("sequential")
## inspect the predictorMatrix
pm_df <- data$mult_imp$predictorMatrix |>
  as.data.frame()
  
rownames(pm_df) <- colnames(pm_df)

library(pheatmap)
pheatmap(pm_df, cluster_rows = FALSE, cluster_cols = FALSE)
```


## Running multiple regressions

```{r, pooled_imp_results, eval=FALSE}
reg_multimp <- with(data = data$mult_imp, lm(pain_intensity ~ sex * age))

summary(reg_multimp$analyses[[1]])
plot(reg_multimp$analyses[[1]])

#pool
reg_pooled <- pool(reg_multimp)
#analyse pooled results - does the confidence interval include both directions? 
summary(reg_pooled, conf.int = TRUE, conf.level = 0.95) |> 
  as_tibble() |>
  janitor::clean_names() |>
  dplyr::filter(term != "(Intercept)") |>
  ggplot(
    aes(x = term,
        y = estimate)
  ) +
  geom_point(colour = "red", size = 2) +
  geom_errorbar(aes(ymin=estimate-x2_5_percent, ymax=estimate+x97_5_percent), width=.2,
                 position=position_dodge(.9)) +
  ylim(c(-0.5, 0.5)) +
  theme_bw()

```

## Complete the dataset with this imputation
```{r, confirm_imputed}
#create imputed data to work with
data$data_imputed <- complete(data$mult_imp, 1) |>
  as_tibble()
  
naniar::vis_miss(data$data_imputed)

## add atttributes to imputed dataframe
attr(data$data_imputed, "meta") <- "Imputed data. Input data is data variable subset from data [file: './data/df_non_imp_select.rds']. Dataframe stored on disk as [file: './data/df_imp_select.rds']"  

## remove clutter
rm(df_imp, df_multimp, reg_multimp, reg_pooled, pm_df, pred, var_selection, list_histograms)

```

## Visualize the imputations
```{r, visualize_imputations, eval=FALSE}
# Load the necessary libraries
library(ggplot2)

# Assuming 'data_with_missing' is your dataset with missing values
# and 'data_with_imputed' is your dataset with imputed values

## dummies
data_missing = data$data_preprocessed
data_imputed = data$data_imputed
#variable_name = names(data$data_preprocessed)[33]


## raw data plot
# Plot for dataset with imputed values
plot_missingness_points <- function(df, var){
  df |>
    ggplot(aes(x = var, y = value)) +
    geom_point(aes(colour = is_missing), position = "jitter", shape = 1) +
    ggtitle(var)
}

#df = joined - distro function
plot_missingness_distro <- function(df, var){
  df |>
    ggplot(aes(x = value)) +
    geom_density(aes(colour = var)) +
    ggtitle(var) +
    facet_wrap(~var)
}

#plot_missingness_distro(df = df)

compare_datasets <- function(
    variable_name,
    data_missing, 
    data_imputed, 
    plot_function) {
  # Create a bar plot to compare the distribution of the specified variable
  
  # Plot for dataset with missing values
  joined <- tibble(
    missing = data_missing[[variable_name]],
    imputed = data_imputed[[variable_name]]
    
  ) |>
    mutate(
      is_missing = is.na(missing)
    ) |>
    pivot_longer(
      c(missing, imputed), names_to = "var", values_to = "value")
    
  plot_missingness <- plot_function(joined, var = variable_name)
    
  return(plot_missingness)
  
}

var_names_preprocessed <- names(data$data_preprocessed) 
var_names_imputed <- names(data$data_imputed)
var_names_iteration <- var_names_preprocessed

map(
  .x = var_names_iteration,
  .f = compare_datasets,
  data_missing = data$data_preprocessed,
  data_imputed = data$data_imputed,
  plot_function = plot_missingness_points
) -> plot_points_missingness_list


names(plot_points_missingness_list) <- var_names_iteration
plot_points_missingness_list

map(
  .x = var_names_iteration,
  .f = compare_datasets,
  data_missing = data$data_preprocessed,
  data_imputed = data$data_imputed,
  plot_function = plot_missingness_distro
) -> plot_distros_missingness_list


names(plot_distros_missingness_list) <- var_names_iteration
plot_distros_missingness_list


```

## Skimming the data
```{r, skimr}
data$data_imputed %>% 
  skimr::skim(
    pain_intensity, 
    sex, 
    age, 
    painint_6weeks, 
    alcohol,
    painint_3months, 
    painint_6months
    ) 
```


## Add `attitude`
We decided that the 'attitude' variable could not be imputed, so we removed that variable from the data and now we put it back. It is the only variable in the data that has missingness now.
```{r}
data$data_imputed$attitude <- data$data_raw$Attitude
# TODO: create merge based on patient code
data$data_imputed <- data$data_imputed |> sjlabelled::remove_all_labels()
data_final <- data$data_imputed

naniar::vis_miss(data$data_imputed)
```

## Save to disk
```{r}
data |>
  write_rds(here::here("data", "data-list.rds"))

write_rds(data$data_imputed, here::here("data", "df_imputed.rds"))
```
