---
params:
  cutoff_painint_high: 3
  cutoff_chronic: 9
---

# Statistical exploration

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  error = FALSE,
  message = FALSE,
  warning = FALSE
  )

source(
  here::here(
    "R",
    "plot_distro.R"
  )
)
```

## Packages
```{r}
library(tidyverse)
library(gt)
library(webshot2)

```

## Data
The data is loaded from disk. 
```{r}
## non-imputed
df_imp <- read_rds(here::here("data", "df_imputed.rds"))
data_list <- read_rds(here::here("data", "data-list.rds"))
df_non_imp <- data_list$data_preprocessed

## add patient code
df_non_imp$patient_code <- data_list$data_subset$patient_code
df_imp$patient_code <- data_list$data_subset$patient_code
```

## Global parameters
```{r}
cutoff_chronic <- params$cutoff_chronic
cutoff_painint_high <- params$cutoff_painint_high
```

## Statistical analysis methods and missing data

### Distribution data and Table 1

For table 1 and the following analysis we have to create a new
    variable based on the variables: `painint_6weeks`, `painint_3months`
    and `painint_6months`. The new variable name: `is_painint_chronic`, with

 -   the outcome no = FALSE and
 -   yes = TRUE. (FALSE = no chronic pain, TRUE = chronic pain).
 -   TRUE = chronic pain = if the variables `painint_6weeks`,
    `painint_3months` and `painint_6months` are each 3 or higher

```{r}
## new variable `chronic_pain` & recode to dichotomous
df_imp <- df_imp |>
  dplyr::mutate(
    painint_6weeks_high = ifelse(painint_6weeks >= cutoff_painint_high, TRUE, FALSE),
    painint_3months_high = ifelse(painint_3months >= cutoff_painint_high, TRUE, FALSE),
    painint_6months_high = ifelse(painint_6months >= cutoff_painint_high, TRUE, FALSE)
)

df_imp <- df_imp |>
  rowwise() |>
  mutate(
    painint_high = 
      sum(painint_6weeks_high, painint_3months_high, painint_6months_high)
  ) 


df_imp <- df_imp |>
  rowwise() |>
  mutate(
    painint_total_score = 
      sum(painint_6weeks, painint_3months, painint_6months)
  ) 

df_imp <- df_imp |>
  mutate(
    is_painint_chronic = ifelse(
      painint_high >= 3, TRUE, FALSE
    )
  )

```

## Reformat dataframe to stacked format
```{r}
df_imp_long <- df_imp |>
  pivot_longer(
    painint_6weeks:painint_6months,
    values_to = "painint_score",
    names_to = "painint_time"
  )


```

## Create time variable
```{r}
## seperate 
df_imp_long <- df_imp_long |>
  separate_wider_delim(painint_time, delim = "_", names = c(NA, "time"))

```

## Recode time
```{r}
## recode to days
df_imp_long <- df_imp_long |>
  mutate(
    time_days = recode_factor(
           time, 
           `6weeks` = 42,
           `3months` = 90,
           `6months` = 180)) 

## write to disk
write_rds(df_imp, file = here::here("data", "imputed_with_attitude_and_chronic.rds"))
```

```{r}
## add var to trace missingness
df_imp_long <- df_imp_long |>
  mutate(is_chronic_missing = is.na(painint_total_score))

## check with visual
df_imp_long |>
  ggplot(aes(
    x = painint_total_score,
    y = as_factor(is_painint_chronic)
  )) +
  geom_point(aes(colour = as_factor(is_painint_chronic)), 
             position = "jitter", shape = 1, alpha = 0.5) +
  scale_color_manual(values = c("darkblue", "magenta")) +
  geom_vline(
    xintercept = cutoff_chronic-1, 
    linetype = "dashed", 
    colour = "darkred", 
    linewidth = 1
    ) +
  guides(color = guide_legend(override.aes = list(size = 3, shape = 1))) +
  xlab("Total painint score") +
  ylab("Recoded to chronic pain") +
  ggtitle("Chronic pain") +
  xlim(c(-2, 30)) -> plot_chronic
plot_chronic

```

## Adding baseline as time=0
```{r}
df_imp_long |>
  group_by(patient_code) |>
  reframe(painint_score = unique(pain_intensity)) |>
  mutate(time_days = 0,
         time_days = as_factor(time_days)) -> df_baseline


df_baseline

df_imp_long <- full_join(df_imp_long, df_baseline) |>
  relocate(is_chronic_missing)
```

## Carry forward
Due to the join we have missing values. We can use a carry forward for the last set of observations and fill in the NA for that value. We group by patient code.
```{r}
df_imp_long |> 
  arrange(patient_code) |>
  tidyr::fill(is_chronic_missing:is_painint_chronic) -> df_imp_long

## add 'baseling' level to 'time'
df_imp_long <- df_imp_long |>
  mutate(time = as.character(time)) |>
  mutate(time = ifelse(is.na(time), "baseline", time))

```

```{r}
## get patient ids
p_ids <- df_imp_long$patient_code |> unique()

## time has wrong level order
df_imp_long$time_days |>
  levels()

## reoder levels
df_imp_long <- df_imp_long |>
  mutate(
    time_days_relevel = fct_relevel(
      time_days, "0", "42", "90", "180"
    ))

## pain over time
df_imp_long |>
  ggplot(
    aes(x = time_days_relevel,
        y = painint_score)) +
  geom_point(aes(
    colour = as_factor(patient_code)), 
    show.legend = FALSE, 
    position = "jitter") +
  xlab("Time (days)") +
  ylab("Pain intensity score") +
  facet_wrap(~is_painint_chronic) -> pain_over_time_chronic

pain_over_time_chronic + citrulliner::theme_individual()

ggsave(here::here("img", "pain_over_time.png"),
       height = 20, width = 30, units = "cm", dpi = 300) 

## calculate average trend
df_imp_long_summary <- df_imp_long |>
  group_by(time_days_relevel, is_painint_chronic) |>
  summarise(median_painint_score = median(painint_score))

df_imp_long_summary

df_imp_long |>
ggplot(
    aes(x = time_days_relevel,
        y = painint_score)) +
  geom_point(aes(
    colour = as_factor(patient_code)), 
    show.legend = FALSE, 
    position = "jitter") +
  xlab("Time (days)") +
  ylab("Pain intensity score") +
  geom_line(
    data = df_imp_long_summary, 
    aes(x = time_days_relevel, 
        y = median_painint_score, 
        group = is_painint_chronic),
    linewidth = 2) +
  facet_wrap(~is_painint_chronic) -> painint_over_time_with_trend


ggsave(plot = painint_over_time_with_trend + citrulliner::theme_individual(),
  here::here("img", "pain_over_time_trend.png"),
       height = 20, width = 20, units = "cm", dpi = 300) 

```

## Rework the graph above to get cummulative pain intensity scores
```{r}
df_imp_pain_cumsum <- df_imp_long |>
   group_by(patient_code) %>%
  arrange(time_days_relevel) |>
  nest()

df <- df_imp_pain_cumsum$data[[1]]
df
## get cumsum (into function)
get_pain_cumsum <- function(df){
  df <- df |>
    arrange(time_days_relevel) |>
  mutate(pain_cumsum = cumsum(painint_score))
  return(df)
  }

## apply function 
df_imp_pain_cumsum |>
  mutate(
    pain_cumsum = map(
      .x = data,
      .f = get_pain_cumsum)
    ) |>
  unnest(pain_cumsum) -> df_cumsum_long

## plot
df_cumsum_long |>
  ggplot(
    aes(x = time_days_relevel,
        y = pain_cumsum)) +
 # geom_line(aes(group = patient_code, colour = as_factor(patient_code)), show.legend = FALSE)  +
  geom_point(aes(colour = as_factor(patient_code)), show.legend = FALSE, position = "jitter") +
  xlab("Time (days)") +
  ylab("Cummulative pain intensity score") +
  facet_wrap(~is_painint_chronic) -> pain_over_time_chronic_cumsum

pain_over_time_chronic_cumsum

ggsave(plot = pain_over_time_chronic_cumsum, filename = here::here("img", "pain_over_time_cumsums.png"),
       height = 50, width = 50, units = "cm", dpi = 300) 

## calculate average trend
df_summary_cumsum <- df_cumsum_long |>
  group_by(time_days_relevel, is_painint_chronic) |>
  summarise(median_painint_score_cumsums = median(pain_cumsum))

pain_over_time_chronic_cumsum +
  geom_point(
    data = df_summary_cumsum, 
      aes(x = time_days_relevel, y = median_painint_score_cumsums), 
      colour = "darkred", size = 6) -> pain_cumsums_over_time

pain_cumsums_over_time

ggsave(plot = pain_cumsums_over_time, filename = here::here("img", "pain_over_time_cumsums_trend.png"),
       height = 20, width = 20, units = "cm", dpi = 300) 

```

## Get individual lines for each patient
```{r}
## create two panels, one with chonic pain and one without.

df_imp_long |>
  dplyr::filter(patient_code %in% c(101:201)) |>
ggplot(
    aes(x = time_days_relevel,
        y = painint_score)) +
  geom_point(aes(colour = is_painint_chronic), 
    show.legend = TRUE,
    size = 3) +
  xlab("Time (days)") +
  ylab("Pain intensity score") +
  geom_line( 
    aes(x = time_days_relevel, 
        y = painint_score, 
        group = is_painint_chronic,
        colour = is_painint_chronic),
    linewidth = 2,
    show.legend = TRUE) +
  facet_wrap(~ patient_code) + 
  ylim(-1, 10) +
  citrulliner::theme_individual() -> painint_over_time_with_trend_individual

painint_over_time_with_trend_individual

painint_over_time_with_trend_individual |>
  ggsave(filename = here::here("img", "pain_over_time_trend_individual.png"),
       height = 70, width = 70, units = "cm", dpi = 300) 
```

## Write to disk as excel file and .rds R binary file.
```{r}
df_imp_long <- df_imp_long |>
  sjlabelled::remove_all_labels()


openxlsx::write.xlsx(
  df_imp_long, 
  file = here::here(
    "data",
    "df_imp_inferential_stats.xlsx"
  )
)

readr::write_rds(
  df_imp_long, 
  file = here::here(
    "app",
    "df_imp_long.rds"
  )
)
```

## Distribution of the data

For all variables at baseline and the outcome variables
'Painint_6weeks', 'Painint_3months' and 'Painint_6months' and
'chronic_pain'. For the outcome variable chronic_pain the distribution
difference between Man/Female, Pain_intensity (baseline) and Age.

```{r}

## variable distributions
df_complete <- map_df(
  .x = df_imp_long,
  .f = as.numeric
)

distros <- map2(
  .x = df_complete,
  .y = names(df_complete),
  .f = plot_distro
)

cowplot::plot_grid(plotlist = distros)

## chronic pain over sex
df_complete |>
  group_by(sex, is_painint_chronic) |>
  tally() |>
  ggplot(aes(
    x = as_factor(sex),
    y = n
  )) +
  geom_col(aes(fill = as_factor(sex)), position = "dodge", show.legend = FALSE) +
  facet_wrap(~as_factor(is_painint_chronic)) +
  xlab("Gender") +
  ylab("Count") +
  ggtitle("Distrubution of gender over people\nwith and without chronic pain")

## pain intensity over age
df_imp_long |>
  ggplot(aes(
    x = age,
    y = pain_intensity
  )) + geom_point(aes(colour = as_factor(sex)), position = "jitter", shape = 1) #+
#  facet_wrap(~pain_intensity)
```

## Table: baseline characteristics of the included patients 
The candidate prognostic factors and the outcome variable chronic pain of patients with acute- or subacute nonspecific idiopathic, non-traumatic neck pain.

```{r}
## function to calculate percentage missing of vector x

missingness_perc <- function(x){
  
  missingness_sum <- sum(is.na(x))
  length_x <- length(x)
  
  perc <- (missingness_sum/length_x)*100
  return(perc)
}

df_non_imp |>
  sjlabelled::remove_all_labels() |>
  group_by(
    sex
  ) |>
  summarise(
    mean_age = mean(age, na.rm = TRUE),
    median_age = median(age, na.rm = TRUE),
    mean_bmi = mean(bmi, na.rm = TRUE),
    median_pain_6_weeks = median(painint_6weeks, na.rm = TRUE),
    missingness_6_weeks = missingness_perc(painint_6weeks),
    median_pain_3_months = median(painint_3months, na.rm = TRUE),
    missingness_3_months = missingness_perc(painint_3months),
    median_pain_6_months = median(painint_6months, na.rm = TRUE),
    missingness_6_months = missingness_perc(painint_6months)
    ) -> df_table_1

df_table_1

df_non_imp |>
  group_by(sex, smoking, alcohol) |>
  tally() -> df_table_2

df_table_2
```

This table includes also the variable pain intensity at 6 weeks, 3
months and 6 months. The outcome variable chronic pain is based on these
three factors.
 Including the percentage (%) of the missing values.
 For the continuous variables a mean (SD) and median.
 For the continue variables I want to add an explanatory note with the scale and the abbreviations in the table. Is this possible to do this directly in R?

To generate tables directly from R, see: https://gt.rstudio.com/

```{r}

df_table_1 |>
  gt() |>
  tab_header(
    title = "Study subjects information",
    subtitle = "Patient details summary") -> table_1_markup

gtsave(table_1_markup, filename = here::here("data", "table1.html"))
gtsave(table_1_markup, filename = here::here("data", "table1.docx"))


df_table_2 |>
  gt() |>
  tab_header(
    title = "Smoking and alcohol use",
    subtitle = "Patient details summary") -> table_2_markup

gtsave(table_2_markup, filename = here::here("data", "table2.html"))
gtsave(table_2_markup, filename = here::here("data", "table2.docx"))

```

## Testing assumptions before backward analysis

1.  For the development of a prediction model and internal validation
    (and testing for assumptions). You need the following packages:
2.  We have to test the assumption for applying a backward analysis.
3.  Linearity, check whether the relationships between the predictor
    variables and the outcome variable are linear. (assessed through
    techniques like scatter plots, partial regression plots, or residual
    plots).

### Scatter plot

For example
```{r}
df_imp_long |> 
  ggplot(aes(
    x = age,
    y = painint_score
  )) +
  geom_point(aes(colour = as_factor(time_days)))
  

```

