---
title: "Chronic neck pain - Data Analysis"
author: "Marc A.T. Teunis & Martine Verwoerd"
site: bookdown::bookdown_site
output: 
    bookdown::htm_book:
        theme: united
        css: style.css
        number_sections: true
        anchor_sections: false
        split_by: chapter
        config:
            sharing:
                 github: yes
                 facebook: no
                 twitter: no
                 all: no
            toc:
                collapse: section
                scroll_highlight: yes
                before: <li class="toc-logo"><a href="./"></a> <h4 class=".paddingtitel ">painr</h2></li>
header-includes:
  - \usepackage{fontawesome5}
---

```{r include = FALSE}
# Define the web addresses for the course
# run as needed, not included automatically because of stuff (aka git complains)
githuburl <- "https://github.com/uashogeschoolutrecht/painr"
readerurl <- "https://rstudio-connect.hu.nl/painr/"
email <- "marc.teunis@hu.nl"

save(
    list = c("githuburl", "readerurl", "email"), 
    file = "course_urls.RData")
```

```{r include = FALSE, eval = TRUE}
# set CSS for objects
knitr::opts_chunk$set(
  class.source="Rchunk", 
  class.output="Rout", 
  warning = FALSE,
  error = FALSE,
  message = FALSE)
```

# Introduction {-}

General intro to the project and the data

Download the source code [`r icons::fontawesome("github")`](https://github.com/uashogeschoolutrecht/painr)

**Martine -> please write a general introduction on the project here: --->
...

## Data Flow
The following diagram discribes how data files and Rmd scripts are connected. Raw data input file lives in `./data-raw`, data output files are written to `./data` and the `.Rmd` files live in ./Rmd.

```{r}
source(
  here::here(
    "inst",
    "data_flow.R"
  )
)
data_flow
```

Please also provide attribution to R itself
```{r}
citation()
```

The `{tidyvese}`
```{r}
citation(package = "tidyverse")
```

`{tidymodels}`
```{r}
citation(package = "tidymodels")
```

